Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-03-23 paper/paper.tex
thorough proofread 6 (sec 5.2 appendix C) * almost done with mnist: awfully many incorrect and invalid observations were made * compacted the text in mnist-like intro and split mnist-like figure in two * reduced the volume of text in the mnist-related appendix * reworded the cplx-netowrk section * moved compression-threshold figure to musicnet * replaced textwidth with columnwidth

Complex networks are, in general, non-holomorphic, i.e. not $\cplx$-differentiable, which is
exacerbated by the loss being real-valued. Indeed a $\cplx$-differentiable real-valued function is
necessarily trivial by Cauchy-Riemann conditions. This issue is dealt with by employing Wirtinger,
or $\cplx\real$ calculus, which generalizes holomorphic calculus to non-holomorphic functions
of $\cplx$-argument \citep{adali_complex-valued_2011,boeddeker_computation_2019}. It satisfies
the product and chain rules, respects complex conjugation and linearity, but most importantly the
$\cplx\real$-differential of a function on $\cplx$ coincides with its classical differential as
a function on $\real^2$. Essentially, it allows straightforward retrofiting of $\cplx$-valued
networks into existing $\real$ deep learning auto-differentiation frameworks. It was considered
as a basis for $\cplx$ version of back-propagation by \citet{benvenuto_complex_1992}, (see
appendix~\ref{sub:wirtinger_calculus}).
% \todo{read on holomorphic nets}

@@

Consider the following parameterization of $W$: $
  W_{ij} = \mu_{ij} \xi_{ij}
$, $\xi_{ij} \in \real$ with $
  \xi_{ij} \sim \mathcal{N}(1, \alpha_{ij})
$, yet $\mu \in \cplx^{n \times m}$. This case corresponds to inference regarding
relevance multipliers $\xi$ rather than the parameters themselves. Under this parameterization
the weight distribution is fully factorized degenerate $\cplx$-Gaussian approximation
\eqref{eq:c-gauss-vi-general} with $
  \Sigma_{ij} = \alpha_{ij} \lvert \mu_{ij} \rvert^2
$ and $
  C_{ij} = \alpha_{ij} \mu_{ij}^2
$. Unlike proper $\cplx$-Gaussian approximation, the complex relation parameter in
\eqref{eq:cplx-gauss-trick} is nonzero in this case and equals $
  \sum_j \alpha_{ij} (x_{ij} \mu_{ij})^2
$.
%
The KL-divergence term for this approximation coincides with \eqref{eq:improper-kl-div-real},
however the major drawback of this approximation is that the additive noise reparameterization,
which disentangles the gradient with respect to $\mu$ from the local output noise, cannot be
applied.

@@

We study the compression-performance trade-off of the proposed Bayesian sparsification
method for $\cplx$-valued networks presented above on image classification and music
annotation tasks.
%
Following \citet{wolter_complex_2018} and \citet{trabelsi_deep_2017}, the class logit scores
are taken as real part of the complex-valued output of a network.
%
% image features
Since image data is not naturally $\cplx$-valued, we preprocess it by either embedding
$\real$-data into $\cplx$ assuming $\Im z = 0$ (\texttt{raw}), or applying the two-dimensional
Fourier Transform (\texttt{fft}), shifting lower frequencies to the centre of the image.
We do not investigate the option to train an auxiliary model to synthesize the imaginary
component from the real data, proposed by \citet{trabelsi_deep_2017}.

@@

Each network within every experiment passes through three successive stages of training:
pre-training, sparsification, and fine-tuning (sec.~\ref{sub:fitting_and_sparsification}).
%
Unless specified otherwise, every experiment uses global gradient clipping within an
$\ell_2$-norm ball of radius $0.5$ and uses ADAM optimizer with the base learning rate
of $10^{-3}$.

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is computed based on the number of IEEE754 floating point values
needed to store the model and equals $
  \tfrac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros, $n_\mathtt{par}$ is the total
number of floats. In particular, each parameter is counted once in a $\real$-valued network,
but twice in a $\cplx$-network. Each model has a compression limit, determined by its
layers' biases, shift and scaling from $\real$- and $\cplx$-valued batch normalization.

\subsection{Stagewise training} % (fold)
\label{sub:fitting_and_sparsification}

@@

\subsection{Metrics and compression} % (fold)
\label{sub:metrics_and_compression}

% perfromance metrics
We use accuracy as the performance metric for multi-class image classification on MNIST-like
and CIFAR10 datasets. Similarly to \citet{trabelsi_deep_2017} performance in experiments with
MusicNet is measured with (pooled) average precision. The evaluation is done after ``dense''
stage, before the beginning and upon termination of the ``fine-tune'' stage.

% subsection metrics_and_compression (end)

@@

We conduct a moderately sized experiment on MNIST and similar $28\times 28$ greyscale image
datasets to study the compression-performance trade-off of the proposed $\cplx$-variational
dropout.
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks and gauge the final performance on the usual test split.
Limiting the size was done primarily to test the regularization strength of compression.

% model architectures
% discussion of TwoLayerDenseModel (TLDM) and SimpleConvModel (SCM)
The $\real$/$\cplx$ networks have been chosen for the purpose of illustrating the compression
and understanding the effects of experiment parameters. \textit{SimpleConvModel} (SCM) is a $2d$
ReLU convolutional network with average pooling ($2\times 2$) after non-linearities with
two convolutions (kernel $5\times 5$, stride $1$) with filters $n_1 \to n_2$ followed by
dense layers $n_3 \to n_\mathtt{out}$. The second network, \textit{TwoLayerDenseModel} (TLDM),
is a wide dense network $n_1 \to n_\mathtt{out}$. For SCM model $n_1, n_2, n_3$ are $20, 50, 500$,
respectively, and for TLDM model $n_1 = 4096$ regardless of their type ($\real$ or $\cplx$),
although \citet{monning_evaluation_2018} argue that this renders models incomparable.

% first experiment: 40-75-40 with fast scheduler
In this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base
learning rate of ${10}^{-3}$ is reduced after the $10$-th epoch to ${10}^{-4}$.
%
Plots~\ref{fig:mnist-like__trade-off} depict samples from the performance-compression curve
on KMNIST and EMNIST-Letters datasets for the studied models and VD compression method
(sec.~\ref{ssub:vd_prior}). Besides that we try \texttt{raw} or \texttt{fft}
input features, vary $
  C \in \{\tfrac3{2^{k+1}}\colon k=1..16\}
$, and repeat the experiment $5$ times.

Shaded horizontal bands reflect the spread of the performance of the uncompressed network
after ``dense'' stage. Each point represents a precision-compression value attained by the
network after ``fine-tune'' stage. The end of each point's tail represents the performance
gain or loss from running the ``fine-tune'' stage.
%
We make a couple of observations based on the results of these plots. First, \texttt{fft}
features perform on par with raw colour channel data in the convolutional networks across most
compression rates, but require high compression rate to make wide shallow dense network work
as well. The upward trend manifests itself in MNIST and Fashion MNIST experiments as well
(see appendix~\ref{sec:mnist_like_experiments}). Finally, overall $\cplx$ variational dropout
methods seem to compress networks adequately well without much loss in performance.

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/legacy__VD__fashionmnist__fft__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{figure__mnist-like__trade-off/legacy__VD__kmnist__fft__-0.5.pdf}
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/legacy__VD__fashionmnist__raw__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{figure__mnist-like__trade-off/legacy__VD__kmnist__raw__-0.5.pdf}
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/VD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/legacy__VD__mnist__fft__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{figure__mnist-like__trade-off/legacy__VD__emnist_letters__fft__-0.5.pdf}
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/legacy__VD__mnist__raw__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{figure__mnist-like__trade-off/legacy__VD__emnist_letters__raw__-0.5.pdf}
    % \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/VD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy trade-off on KMNIST and the letters of the Extended MNIST datasets.
    (\textit{top}) \texttt{fft} features, (\textit{bottom}) \texttt{raw} features.
  }
  \label{fig:mnist-like__trade-off}
\end{figure}

@@

@@

@@

To supplement the results, reported in the main text (figures \ref{fig:paper__mnist-like__trade-off__VD__fft}
and \ref{fig:paper__mnist-like__trade-off__VD__raw}), we conduct additional experiments
with the same basic set-up, but over a finer grid of KL-divergence term coefficients.
%
These experiments lend further evidence in support of the conclusions, drawn in the main
paper, namely, that the $\cplx$-valued extension of the $\real$-valued variational dropout
can offer significant compression levels without much loss in performance. At the same
time, the test performance of both $\real$ and $\cplx$ compressed networks benefits from
an extra fine-tuning stage, during which sparsity masks are fixed.

One intriguing observation, is that on MNIST-like images the dense network, taking Fourier
features on input, exhibits gradually improving test-split performance. This is in stark
contrast to raw image input, where the performance stays at one level in the small compression
regime, and then gradually decays at high compression levels. Noteworthy is the presence
of a trough in the pre fine-tuning stage performance in the small compression regime on figures
\ref{fig:appendix__mnist-like__trade-off__ARD__raw},
\ref{fig:appendix__mnist-like__trade-off__VD__raw},
\ref{fig:appendix__cmp__mnist-like__trade-off__ARD__raw},
and \ref{fig:appendix__cmp__mnist-like__trade-off__VD__raw}, which then slowly disappears
as compression increases.

The following scatter plots depict the samples from the compression-performance trade-off
curve of simple networks studied in the experiments on the MNIST-like datasets. Transparent
horizontal bands on each plot represent min-max performance spread of an uncompressed
network on the test split. The points illustrate the final trade-off of the compressed
networks after fine-tuning, while their tails show the performance gain or loss due to
the fine-tuning.

We recap the set-up of each additional experiment on MNIST-like datasets:
\begin{itemize}
  \item staged training, with weight and optimizer state transfer
  \item $40$, $75$, and $40$ epochs for ``dense'', ``sparsify'', and ``fine-tune'' stagers, respectively
  \item ADAM optimizer with learning rate $10^{-3}$ before the $10$-th, and $10^{-4}$
  afterwards
  \item global $\ell_2$-norm gradient clipping set at $\tfrac12$
  \item training batches of size $128$ from a fixed randomly subset ($10^4$ images) of the training split
  \item the multiplier of the Kullback-Leibler term in ELBO varies over $
    C \in \{
      \tfrac32 2^{-\tfrac{k}2} \colon k=2, \cdots, 38
    \}
  $
  \item the sparsification threshold $\tau$ is set to $-\tfrac12$
\end{itemize}
Each experiment has been performed $5$ times to obtain a small sample of performance-compression
values for each setting of $C$.

@@

The following observations can be made from the conducted experiment:
\begin{itemize}
  \item $\cplx$-valued networks do not have as much spare capacity as $2 \real$ due to
  constants imposed by multiplication in $\cplx$ numbers (\ref{fig:appendix__cmp__mnist-like__trade-off__ARD__raw},
  \ref{fig:appendix__cmp__mnist-like__trade-off__VD__raw})
  \item Variational Dropout and Automatic Relevance Determination offer similar compression-performance
  trade-off, with negligibly lower compression by the latter (both $\real$ and $\cplx$ versions,
  e.g. figures \ref{fig:appendix__mnist-like__trade-off__ARD__raw} and \ref{fig:appendix__mnist-like__trade-off__VD__raw})
  \item smaller networks tend to be less compressible (figures \ref{fig:appendix__cmp__mnist-like__trade-off__ARD__fft},
  \ref{fig:appendix__cmp__mnist-like__trade-off__VD__fft})
\end{itemize}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__raw}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__raw}
\end{figure}

>>>>

Complex-valued networks are, in general, non-holomorphic, i.e. not $\cplx$-differentiable, which is
exacerbated by the loss being real-valued. Indeed a $\cplx$-differentiable real-valued function is
necessarily trivial by Cauchy-Riemann conditions. This issue is dealt with by employing Wirtinger,
or $\cplx\real$ calculus, which generalizes holomorphic calculus to non-holomorphic functions
of $\cplx$-argument \citep{adali_complex-valued_2011,boeddeker_computation_2019}. It satisfies
the product and chain rules, respects complex conjugation and linearity, but most importantly the
$\cplx\real$-differential of a function on $\cplx$ coincides with its classical differential as
a function on $\real^2$. Essentially, it allows straightforward retrofiting of $\cplx$-valued
networks into existing $\real$ deep learning auto-differentiation frameworks. It was considered
as a basis for $\cplx$ version of back-propagation by \citet{benvenuto_complex_1992}, (see
appendix~\ref{sub:wirtinger_calculus}).
% \todo{read on holomorphic nets}

@@

Consider the following parameterization of $W$: $
  W_{ij} = \mu_{ij} \xi_{ij}
$, $\xi_{ij} \in \real$ with $
  \xi_{ij} \sim \mathcal{N}(1, \alpha_{ij})
$, yet $\mu \in \cplx^{n \times m}$. This case corresponds to inference regarding
relevance multipliers $\xi$ rather than the parameters themselves. Under this parameterization
the weight distribution is fully factorized degenerate $\cplx$-Gaussian approximation
\eqref{eq:c-gauss-vi-general} with $
  \Sigma_{ij} = \alpha_{ij} \lvert \mu_{ij} \rvert^2
$ and $
  C_{ij} = \alpha_{ij} \mu_{ij}^2
$. Unlike proper $\cplx$-Gaussian approximation, the complex relation parameter in
\eqref{eq:cplx-gauss-trick} is nonzero in this case and equals $
  \sum_j \alpha_{ij} (x_{ij} \mu_{ij})^2
$.
%
The KL-divergence term coincides with \eqref{eq:improper-kl-div-real}, however the major
drawback of this approximation is that the gradient of the loss with respect to $\mu$ cannot
be disentangled from the local output noise by additive reparameterization.

@@

We study the compression-performance trade-off of the proposed Bayesian sparsification
method for $\cplx$-valued networks presented above on image classification and music
annotation tasks.
%
Following \citet{wolter_complex_2018} and \citet{trabelsi_deep_2017}, the class logit scores
are taken as real part of the complex-valued output of a network.
%
% image features
Since image data is not naturally $\cplx$-valued, we preprocess it by either embedding
$\real$-data into $\cplx$ assuming $\Im z = 0$ (\texttt{raw}), or applying the two-dimensional
Fourier Transform (\texttt{fft}), centering the lower frequencies. We do not train an
auxiliary model to synthesize the imaginary component from the real data, proposed by
\citet{trabelsi_deep_2017}.

@@

Each network within every experiment passes through three successive stages of training:
pre-training (\textit{``dense''}), sparsification (\textit{``sparsify''}), and fine-tuning
(\textit{``fine-tune''}) (sec.~\ref{sub:staging}). The parameters and optimizer state are
passed between stages, except for last one, which inherits just the parameters.
%
Every experiment uses ADAM optimizer, resets the learning rate to $10^{-3}$ before each
stage, and applies global $\ell_2$-norm gradient clipping at $0.5$.

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is calculated based on the number of IEEE754 floating point values
needed to store the model and equals $
  \tfrac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros, $n_\mathtt{par}$ is the total
number of floats. In particular, each parameter is counted once in a $\real$-valued network,
but twice in a $\cplx$-network. Each model has a compression limit, determined by biases,
shift and scaling in $\real$- and $\cplx$-valued batch normalization layers.

\subsection{Stagewise training} % (fold)
\label{sub:staging}

@@

@@

We conduct a moderately sized experiment on MNIST-like datasets of $28\times 28$ greyscale
images to study the compression-performance trade-off of the proposed $\cplx$-variational
dropout: MNIST \citep{lecun_gradient-based_1998}, KMNIST \citep{clanuwat_deep_2018},
EMNIST \citep{cohen_emnist_2017} and Fashion-MNIST \citep{xiao_fashion-mnist_2017}.
%
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks and measure the performance with classification \textit{accuracy}
score on the usual test split.
% Limiting the size was done primarily to test the regularization strength of compression
% and to facilitate quicker experiments.

% model architectures
% (fft) C - R - C / 2
% (raw) R - C - 2 * R
We consider two simple architectures for $\real$ and $\cplx$ networks in this experiment,
which have been chosen for the purpose of illustrating the compression and understanding
the effects of experiment parameters. \textit{TwoLayerDenseModel} is a wide dense ReLU
network $784 \to 4096 \to n_\mathtt{out}$, and \textit{SimpleConvModel} is a ReLU net with
two $2d$ $k_5 s_1$ convolutions with filters $20 \to 50$, two $k_2 s_2$ average pooling steps,
and a dense classifier network $800 \to 500 \to n_\mathtt{out}$. The number of intermediate
feature is as specified in $\real$ and $\cplx$ models, is halved in complex models tagged
$\tfrac12\cplx$, and doubled in real models labelled $2\real$.

% experiment setup
Stages (sec.~\ref{sub:staging}) last for $40$, $75$ and $40$ epochs, respectively, in each
experiment. The sparsification threshold $\tau$ is fixed at $-\tfrac12$, the training batch
size is set to $128$ and the base learning rate ${10}^{-3}$ is reduced after the $10$-th epoch
to ${10}^{-4}$ at every stage.
%
% grid specs
We let $
  C \in \{
    \tfrac32 2^{-\tfrac{k}2} \colon k=2, \cdots, 38
  \}
$ in \eqref{eq:elbo_with_coef}, test VD and ARD methods, try \texttt{raw} or \texttt{fft}
input features, and repeat each experiment $5$ times to get a sample of compression-accuracy
pairs.

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__VD__mnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (VD, \texttt{fft}, MNIST):
    $2 \real / \cplx$ (\textit{top}) and $\real / \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__fft}
\end{figure}

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (ARD, \texttt{raw}, MNIST):
    $\real / \cplx$ (\textit{top}) and $\real / \tfrac12 \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__raw}
\end{figure}

% observations
Figures~\ref{fig:mnist-like__trade-off__fft} and~\ref{fig:mnist-like__trade-off__raw}
depict the resulting compression-accuracy trade-off on MNIST for the simple models
described above (other datasets are in appendix~\ref{sec:mnist_like_experiments}).
%
The points represent the trade-off of the compressed networks after fine-tuning, while their
tails illustrate the impact of this stage on the accuracy. Transparent horizontal bands on
each plot represent min-max performance spread of an uncompressed network on the test split.

The overarching pattern in all experiments on MNIST-like datasets is that both ARD and VD yield
compress similarly and ARD offers marginal advantage in terms of performance after fine-tuning.
Fourier features perform marginally worse than raw image data for each model across all
compression rates, and their trade-off for the dense network has a hat-like shape, contrasting
the raw image features, for which the accuracy stays at one level for moderate rates,
and then abruptly decays in high compression regime.

The impact of ``fine-tune'' is mixed, and depends on the dataset and features, although it
frequently improves accuracy in high compression $\times50+$ (high $C$ in \eqref{eq:elbo_with_coef})
for raw image input. On EMNIST Letters and MNIST datasets in the small compression regime the
dense network exhibits a trough in accuracy before fine-tuning stage.

% conclusions
These experiments demonstrate that $\cplx$ Variational Dropout can offer adequate compression
without much loss in performance for $\cplx$-valued networks. Although ARD performs on par
with VD, unlike VD, it uses a proper prior and has a tractable analytic expression for the
KL-divergence \eqref{eq:emp-bayes-kl-div} in both $\real$ and $\cplx$ cases.
% fft: $2\real$ convolutional model trade-off overall better than $\real$, dense -- on par.
% raw: $\cplx$ models trade-off overall better than $\tfrac12\cplx$.
Wider $\real$ and $\cplx$ networks appear to have more favourable trade-off between
compression and accuracy.
% fft: any-$\real$ dense model compresses better than $\cplx$
However, it seems that due to more intrinsic degrees of freedom, $\real$-valued models
perform and compress better than $\cplx$, even when the former uses the same number of
features as the latter.
% $\cplx$-valued networks do not have as much spare capacity as $2 \real$ due to
%  constraints imposed by multiplication in $\cplx$ numbers

@@

We use accuracy as the performance metric for multi-class image classification on CIFAR10 datasets.

@@

% perfromance metrics
Similarly to \citet{trabelsi_deep_2017} performance in experiments with
MusicNet is measured with (pooled) average precision. The evaluation is done after ``dense''
stage, before the beginning and upon termination of the ``fine-tune'' stage.

@@

@@

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__raw}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__raw}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $2\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__cmp__mnist-like__trade-off__ARD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__VD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__VD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__cmp__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $2\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__cmp__mnist-like__trade-off__VD__fft}
\end{figure}
